\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-07 paper.tex
Section on task planning

\subsection{Modeling the interaction}

@@

Experiences comprise of emotions and questions
\subsection{Task planning}

\cite{Alili2008}

@@

We conclude this article we three different perspective on the knowledge-centric
techniques we have presented here: first, the \emph{architect perspective}:
what are the pro and cons of our approach from an engineer point of view. Then,
the \emph{logician perspective}: does the promise of automated reasoning for
robots really comes to fruition with ontologies? And eventually the
\emph{cognitician perspective} or how explicit knowledge manipulation endows
our robots with new cognitive capabilities.

>>>>

\subsection{Modeling the knowledge related to the interaction}

@@

Experiences comprise of emotions and questions

\subsection{Task planning}

In order to devise how a given goal can be accomplished, the robot has to
elaborate a plan,\ie a set of actions to be achieved by the robot and its human
partners.  We use in our architecture the HATP planner~\cite{Alili2009} (for
Human Aware Task Planner).  HATP is based on a Hierarchical Task Network (HTN)
refinement, which performs an iterative task decomposition into sub-tasks until
reaching atomic actions.  The planning domain defines a set of methods
describing how to decompose a task and can be seen as the procedural knowledge
of the robot.

In order to produce a collaborative behaviour, the robot plans not only for
itself but also for the other agents. The planning domain of each agent is
instantiated from the agent-specific model in the ORO server. The resulting
plan, called ``shared plan'' is a set of actions that form a stream for each
agent involved in the goal achievement. HATP can also be tuned by setting up
different costs depending on the actions to apply and by taking into account a
set of constraints called \emph{social rules}. This tuning aims at adapting the
robot's behaviour according to the desired level of cooperation of the robot.

Two important remarks: because HATP is a generic symbolic task planner, we have
been able to design a planning domain at a semantic level which is close to the
one used in the human-robot dialogue (the planner vocabulary contains concepts
like \texttt{give}, \texttt{table}, \texttt{is on}...). Hence only a few
ontology rules have been required to map both the knowledge extracted from the
situation assessment and the statements originated from the verbal interaction
to the planner domain.

Second remark, after some research (see appendix B of~\cite{ma these} for a
detailled discussion)  we have decided not to represent neither tasks nor plans
in the ontology: the planning domain (with tasks pre- and postconditions) is
stored in a specific format, outside of the knowledge repository, and the
resulting plans are directly communicated to the robot controller. The ontology
formalism that we have adopted (namely OWL) does not appear to be a satisfaying
reprepresentation language for concepts that are both tightly bound to time
representation and non-monotonic. Other researches (\cite{tenorth...}) have attempted with more success (\fixme{regarder ca!!}).

@@

We conclude this article with three different perspectives on the knowledge-centric
techniques we have presented here: first, the \emph{architect perspective}:
what are the pro and cons of our approach from an engineer point of view. Then,
the \emph{logician perspective}: does the promise of automated reasoning for
robots really comes to fruition with ontologies? And eventually the
\emph{cognitician perspective} or how explicit knowledge manipulation endows
our robots with new cognitive capabilities.
